Discovering Robustness Threats by Mining the Exception Stack Traces of over 600 Android Projects

### 2014-09-03 android-stacks.tex
addressing comments

\begin{document}
%\title{ What GitHub and Googlecode can tell us about the Causes of Crashes of Android Apps? }
%\title{ A Mining Study the Googlecode can tell us about the Causes of Crashes of Android Apps? }

\title{What can the Issues of GitHub and Googlecode tell us about Exception Fragilities in Android? }

%\title{On the Common Characteristics Android-Apps Crashes: Results of a GitHub and Googlecode Mining Study}

%\title{On the Common Characteristics of Android-Apps Crashes: Results of a GitHub and Googlecode Mining Study}

%\title{On the Common Characteristics of Android Crashes: Results of a GitHub and Googlecode Mining Study}

%\title{A first Glance on Characteristics of Android Crashes: Results of a GitHub and Googlecode Mining Study}

%\title{A first Glance on Characteristics of Android Crashes: Results of a GitHub and Googlecode Mining Study}

@@

%GitHub 315,92
% Googlecode: 127,456
%All: 159,048

\begin{abstract}
The number of mobile apps is increasing in a daily rate. If on the one hand they
extend phones capabilities far beyond of the basic calls, on the other hand they have to 
cope with an increasing number of exceptinal conditions 
(e.g., faults in underlying middleware or hardware; memory and battery restrictions). 
Therefore, mechanisms for exception detection and handling are not an optional add-on but a 
fundamental part of such apps. However, studies have shown that such mechanisms are the less 
understood and tested parts of the system. As a consequence they may affect the system the 
other way around, making the system even less robust.
This paper reports a study which mined the stack traces embedded on 
159,048 issues reported on Android projects hosted in GitHub (482 projs.) 
and Googlecode (157 projs.). The goal of this study was to: identify common characteristics 
of such traces and investigate the fragilities on the exception-related code that can be revealed
from them. Overall 6,005 exception stack traces were extracted, and analyzed in 
combination with source code and bytecode analysis. 
Some fragilities consistently detected in this study include: 
(1) undocumented runtime exceptions thrown by third party code;
(2) undocumented checked exceptions thrown by Android native code;
(3) unexpected exception wrappings (e.g., Errors being wrapped in checked exceptions),
revealing that Java hybrid exception model is not fully used according to its purpose.
Moreover, about 50\% of all analyzed projects reported at least one NullPointerExeption
as the root cause of an exception stack trace.
This study provide important insights on how such fragilities impose challenges 
to application robustness not only in Android environment but also to other 
systems based on Java exception model.

@@

%GitHub 31,592
% Googlecode: 127,456
%All: 159,048

@@

Until recent years, the availability of issues reporting abnormal
computation states (e.g., exception stack traces) was scarce. The
available data comprised only few large scale projects (e.g. the Mozilla
crash report dataset\footnote{https://code.google.com/p/promisedata/issues/detail?id=88}).
Recently we have observed a rapid increase in the number of open
source projects that make their control version as well as the issue tracking systems
publicly available in repository hosting sites such as GitHub and Googlecode.

@@

%482 applications) and Google code (157 applications)

To do so we perform a post mortem analysis of the exception stack traces embedded on all issues
created on Android projects available in GitHub and Googlecode. Using a custom tool called ExceptionMiner,
 which we develop specifically for this study, we mine stack traces from issues across all Android projects on
GitHub and Googlecode. Overall, we analyzed all issues extracted from on 482 Android projects defined in

%GitHub 31,592
% Googlecode: 127,456
%All: 159,048

GitHub and 157 projects defined in Googlecode. Overall it corresponds to 159,048 issues (31,592 issues from GitHub projects 
and 127,456 issues from Googlecode projects) from which  6,005 stack traces were extracted and used as
 a source of information for our analysis. The stack trace analysis was augmented by additional information
extracted from JDK, Android Plaform and the applications using byte code and source code analysis.
 Manual inspection was used to leverage the understanding of exception stack traces and support further discussions and insights.

@@

  \item  It performs a large scale post morten analysis of exception stack traces reported on Android projects hosted in GitHub and Googlecode.

@@

The goal of this mining study was to extract exception stack traces embedded on issues of Android projects 
(hosted in GitHub and GoogleCode), looking for its 
common characteristics and investigating whether
 they could reveal common fragilities on the exception code of both the 
applications and the underlaying  framework.

@@

To answer the research questions the following information is needed: (i) the issues of reported on every Android project hosted on 
GitHub and Googlecode; (ii) the stack traces embedded on those issues; (iii) the type of the exceptions
 reported on the stack traces (e.g., error, runtime, checked); (iv) the origin of such exceptions 
(e.g., the application, a library, the Android platform). Next sections describes how such information 
was retrieved from GitHub and Googlecode, how the exceptions stack traces were extracted and distilled,
 and how the type associated to each exception and its origins was identified.

@@

This set of 482 projects contained overall 32,582 issues from which 4,042 exception stack traces 
were extracted. Issues on Github are different from issues on dedicated bug tracking tools such as 
Bugzilla and Jira. The most important difference is that there are no predefined fields
  (e.g. severity and priority). Instead, Github uses a more open ended tagging system, on which
repositories are offered a pre-defined set of labels, but repository owners can modify 
them at will. Therefore, an issue may have none ore an arbitrary set of labels depending 
on which repository it was created.

@@

\subsubsection{Android Apps in GoogleCode}

Differently from GitHub, Googlecode does not provide an API to access the information related
 to hosted projects \footnote{Google code used to provided a Web service to such repository, but it was deactivated in June 2013 what Google called a "clean-up action".}.

However, Googlecode still contain largely used open-source Android apps (e.g. K9Mail \footnote{K9Mail moved to Github but as a way of not loosing the project history 
it advices their users to report bugs on Googlecode issue tracker.})  and the history of such apps are also valuable. 
To overcome this limitation we needed to implement a Web Crawler (incorporated in Exception Miner tool described next) that navigated 
through the web interface of Googlecode projects extracting all issues and issue comments and storing in a relational database for later analysis.

To identify Android projects in Googlecode, we performed a similar heuristic, we performed a case insensitive search 
(on Googlecode search interface) for the term \textsf{android}.  On January 2014, when we queried Googlecode, such heuristic could
 find 788  projects, from which 724 defined at least 1 issue and, 183 projects 
had at least one issue including a stack trace. Then we performed further cleanup, inspecting the site of every Android
reporting at least one stack trace, and as a result we could identify 157. 
Android projects.  All issues created on such projects (i.e., 41,674 issues) were analyzed and 1,963 stack traces 
were extracted.

%For instance: K9mail  was hosted in Googlecode 
%before moving to GitHub and in order not to loose the project history it advices their 
%users to create issues on Googlecode as a way to preserve project history.

%To overcome this limitation and mine the information available on such Googlecode 
%projects we implemented a Web crawler which navigated (incorporated in Exception Miner tool described next) 
% on each project and issue page, extracting the content of each issue to store on relational database.

@@

\noindent\emph{Repository Connectors.}  This component is responsible to connect to
 the host repositories in order to collect project issues. The GitHub Connector connected
 to GiHub through GHTorrent. The GoogleCode connector comprised of a Web Crawler,
 that traversed Googlecode Web interface in order to extract the project issues.

@@

\subsection{Replication Package}
All the data used in this study will be publicly available at http://.....
Specifically we provide: (i) all issues related to Android projects found
in GITHub and GoogleCode used in this study; (iii) all stack traces extracted
from issues (ii) the results of manual inspection steps; (iv) the tool developed 
in this work to support stack trace extraction and distilling.

@@

    \hline
  \end{tabular}
\caption{Root Exceptions occurrences and popularity in repositories hosted in Googlecode (GC) and GitHub(GH).}
\label{tab:topten}
\end{table*}

@@

In this paper we present an exploratory study in which we mined the stack 
traces embedded in all issues defined in 482 Android projects hosted in Github and 
157 projects hosted in Googlecode.

>>>>

\begin{document}
%\title{ What GitHub and Google Code can tell us about the Causes of Crashes of Android Apps? }
%\title{ A Mining Study the Google Code can tell us about the Causes of Crashes of Android Apps? }

%\title{What can the Issues of GitHub and Google Code tell us about Exception Fragilities in Android? }

%\title{What can the Issues of GitHub and Google Code tell us about Exception Fragilities in Android? }

\title{Discovering Robustness Threats by Mining the Exception Stack Traces of over 600 Android Projects}

%\title{Discovering Fragilities on the Exception-Related Code through  Exception Stack Trace Mining}

%\title{On the Common Characteristics Android-Apps Crashes: Results of a GitHub and Google Code Mining Study}

%\title{On the Common Characteristics of Android-Apps Crashes: Results of a GitHub and Google Code Mining Study}

%\title{On the Common Characteristics of Android Crashes: Results of a GitHub and Google Code Mining Study}

%\title{A first Glance on Characteristics of Android Crashes: Results of a GitHub and Google Code Mining Study}

%\title{A first Glance on Characteristics of Android Crashes: Results of a GitHub and Google Code Mining Study}

@@

%GitHub 315,92
% Google Code: 127,456
%All: 159,048

%The goal of this study was to: identify common characteristics 
%of such traces and investigate the fragilities on the exception-related code that can be revealed
%from them. 
%The goal of this study was to: identify common characteristics 
%of such traces and investigate the whether they can reveal
%the characteristics on the source code that may be favoring the introduction of faults 
%(i.e. Uncaught Exceptions, and Unintened handler Actions), and consequently
%decreasing the program robustness.
%(of apps or the underlaying framework)

\begin{abstract}
The number of mobile apps is increasing in a daily rate. If on the one hand they
extend phones capabilities far beyond of the basic calls, on the other hand they have to 
cope with an increasing number of exceptinal conditions 
(e.g., faults in underlying middleware or hardware; memory and battery restrictions). 
Therefore, mechanisms for exception detection and handling are not an optional add-on but a 
fundamental part of such apps. However, studies have shown that such mechanisms are the less 
understood and tested parts of the system. As a consequence they may affect the system the 
other way around, making the system even less robust, leading to failures such as uncaught exceptions 
and unintended handler actions (i.e., exceptions being mistakenly caught by existing handlers).
This paper reports a study which mined the stack traces embedded on 
159,048 issues reported on Android projects hosted in GitHub (482 projs.) 
and Google Code (157 projs.). 
The goal of this study was to: investigate the whether the exception stack traces
 can reveal common characteristics on the exception-related code that
may be favoring the introduction of the aforementioned failures, which would be responsible for 
decreasing the application robustness. We call such characteristics ``fragilities" of the exception-related code.
Overall 6,005 exception stack traces were extracted, and analyzed in 
combination with source code and bytecode analysis. 
Some fragilities consistently detected in this study include: 
(1) undocumented runtime exceptions thrown by third party code;
(2) undocumented checked exceptions thrown by Android native code;
(3) unexpected exception wrappings (e.g., Errors being wrapped in checked exceptions),
revealing that Java hybrid exception model is not fully used according to its purpose.
Moreover, about 50\% of all analyzed projects reported at least one NullPointerExeption
as the root cause of an exception stack trace.
This study provide important insights on how such fragilities impose challenges 
to application robustness not only in Android environment but also to other 
systems based on Java exception model.

@@

%GitHub 31,592
% Google Code: 127,456
%All: 159,048

@@

Until recent years, the availability of issues reporting abnormal
computation states (e.g., exception stack traces) was scarce. The
available data comprised only few large scale projects (e.g. the Mozilla
crash report dataset\footnote{https://code.google.com/p/promisedata/issues/detail?id=88}).
Recently we have observed a rapid increase in the number of open
source projects that make their control version as well as the issue tracking systems
publicly available in repository hosting sites such as GitHub and Google Code.

@@

%482 applications) and Google code (157 applications)
%GitHub 31,592
% Google Code: 127,456
%All: 159,048

To do so we perform a post mortem analysis of the exception stack traces embedded on all issues
created on Android projects available in GitHub and Google Code. Using a custom tool called ExceptionMiner,
 which we develop specifically for this study, we mine stack traces from issues across all Android projects on
GitHub and Google Code. Overall, we analyzed all issues extracted from on 482 Android projects defined in
GitHub and 157 projects defined in Google Code. Overall it corresponds to 159,048 issues (31,592 issues from GitHub projects 
and 127,456 issues from Google Code projects) from which  6,005 stack traces were extracted and used as
 a source of information for our analysis. The stack trace analysis was augmented by additional information
extracted from JDK, Android Plaform and the applications using byte code and source code analysis.
 Manual inspection was used to leverage the understanding of exception stack traces and support further discussions and insights.

@@

  \item  It performs a large scale post morten analysis of exception stack traces reported on Android projects hosted in GitHub and Google Code.

@@

The goal of this mining study was to extract exception stack traces embedded on issues of Android projects 
(hosted in GitHub and Google Code), looking for its 
common characteristics and investigating whether
 they could reveal common fragilities on the exception code of both the 
applications and the underlaying  framework.

@@

To answer the research questions the following information is needed: (i) the issues of reported on every Android project hosted on 
GitHub and Google Code; (ii) the stack traces embedded on those issues; (iii) the type of the exceptions
 reported on the stack traces (e.g., error, runtime, checked); (iv) the origin of such exceptions 
(e.g., the application, a library, the Android platform). Next sections describes how such information 
was retrieved from GitHub and Google Code, how the exceptions stack traces were extracted and distilled,
 and how the type associated to each exception and its origins was identified.

@@

%GitHub 31,592
% Google Code: 127,456
%All: 159,048

This set of 482 projects contained overall 31,592 issues from which 4,042 exception stack traces 
were extracted. Issues on Github are different from issues on dedicated bug tracking tools such as 
Bugzilla and Jira. The most important difference is that there are no predefined fields
  (e.g. severity and priority). Instead, Github uses a more open ended tagging system, on which
repositories are offered a pre-defined set of labels, but repository owners can modify 
them at will. Therefore, an issue may have none ore an arbitrary set of labels depending 
on which repository it was created. Table ~\ref{tab:lables} illustrates the ocurrences of different lables 
on the issues including exception stack traces.

% \centering
%  \begin{tabular}{|p{2cm}| p{5cm}|}
%    \hline

\begin{table}
  \centering
  \begin{tabular}{lr|lr}
    \hline
     \multicolumn{2}{c}{\bfseries{GitHub}} &  \multicolumn{2}{c}{\bfseries{Google Code}} \\
      \bfseries{Lable} &  \bfseries{\% Occurrences} &  \bfseries{Lable} &  \bfseries{\% Occurrences} \\
    \hline
empty &	54,24\% & Defect &	91,96\% \\
Defect &	39,56\%  & Enhancement  &	3,16\% \\
Enhancement &	0,57\% & Task	& 1,37\% \\
Support &	0,52\% & empty &	1,12\% \\
Problem &	0,36\% & StackTrace &	0,70\% \\
Others &	4,74\% &  Others &	1,68\% \\   
  \hline
  \end{tabular}
  \caption{Lables on issues including exception stack traces.}
  \label{tab:lables}
\end{table}

@@

\subsubsection{Android Apps in Google Code}

Differently from GitHub, Google Code does not provide an API to access the information related
 to hosted projects \footnote{Google code used to provided a Web service to such repository, but it was deactivated in June 2013 what Google called a "clean-up action".}.

However, Google Code still contain largely used open-source Android apps (e.g. K9Mail \footnote{K9Mail moved to Github but as a way of not loosing the project history 
it advices their users to report bugs on Google Code issue tracker.})  and the history of such apps are also valuable. 
To overcome this limitation we needed to implement a Web Crawler (incorporated in Exception Miner tool described next) that navigated 
through the web interface of Google Code projects extracting all issues and issue comments and storing in a relational database for later analysis.

To identify Android projects in Google Code, we performed a similar heuristic, we performed a case insensitive search 
(on Google Code search interface) for the term \textsf{android}.  On January 2014, when we queried Google Code, such heuristic could
 find 788  projects, from which 724 defined at least 1 issue and, 183 projects 
had at least one issue including a stack trace. Then we performed further cleanup, inspecting the site of every Android
reporting at least one stack trace, and as a result we could identify 157. 
Android projects.

This set of 157 projects contained overall 127,456 issues from which 1,963 exception stack traces 
were extracted. Based on the same assumption used for Github dataset selection, we did not retrict 
the analysis for issues labled as Defect.

%SELECT distinct(type), count(*) as n
%  FROM exceptionalissue where repo !='android' group by type order by n desc;

%For instance: K9mail  was hosted in Google Code 
%before moving to GitHub and in order not to loose the project history it advices their 
%users to create issues on Google Code as a way to preserve project history.

%To overcome this limitation and mine the information available on such Google Code 
%projects we implemented a Web crawler which navigated (incorporated in Exception Miner tool described next) 
% on each project and issue page, extracting the content of each issue to store on relational database.

@@

\noindent\emph{Repository Connectors.}  This component is responsible to connect to
 the host repositories in order to collect project issues. The GitHub Connector connected
 to GiHub through GHTorrent. The Google Code connector comprised of a Web Crawler,
 that traversed Google Code Web interface in order to extract the project issues.

@@

\subsection{Replication Package}
All the data used in this study will be publicly available at http://.....
Specifically we provide: (i) all issues related to Android projects found
in GITHub and Google Code used in this study; (iii) all stack traces extracted
from issues (ii) the results of manual inspection steps; (iv) the tool developed 
in this work to support stack trace extraction and distilling.

@@

    \hline
  \end{tabular}
\caption{Root Exceptions occurrences and popularity in repositories hosted in Google Code (GC) and GitHub(GH).}
\label{tab:topten}
\end{table*}

@@

In this paper we present an exploratory study in which we mined the stack 
traces embedded in all issues defined in 482 Android projects hosted in Github and 
157 projects hosted in Google Code.
